# Reproducible Research: Peer Assessment 1

## Loading and preprocessing data

* Data is downloaded if not present .
* Zip file downloaded is uncompressed.
* Data is preprocessed .
* Data is inxpected .

```{r getData, echo=TRUE}

getData <- function() {

    install.packages("downloader",repos="http://cran.rstudio.com/")
    require(downloader)
    url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"  
    
    zip_source_name = "activity.zip"
    
    print("check file")
    
    if(!file.exists(zip_source_name)) {
    
            print("download file")
            download(url,destfile=zip_source_name, mode="wb")          
    }
    
    if (!"act" %in% ls()) { 
                  
            print("unzip file")
            act <- unzip(zip_source_name)
    }   
 
    actData <- read.csv(act, header=T, colClasses = c("integer", "Date", "factor"))
    actData$interval <- factor(actData$interval)
    actData$date <- as.Date(actData$date, format="%Y-%m-%d")
    actData
}
actData <- getData()
```   

Inspect data.
```{r inspectData, echo=TRUE}

head(actData)
dim(actData)
summary(actData)
str(actData)

```

## What is mean total number of steps taken per day?

* Calculate the total number of steps taken per day
* Make a histogram of the total number of steps taken each day
* Calculate and report the mean and median of the total number of steps taken per day

```{r stepsPerDay, echo=TRUE}

 actDataNoNA <- na.omit(actData)
  
 totalStepsPerDay <- aggregate(steps ~ date, actDataNoNA, sum)

 hist(totalStepsPerDay$steps, breaks = 20, main="Histogram of steps ", xlab="Number of steps taken per day", ylab = "Number of times", col="grey")
 
```  
Mean of the total number of steps taken per day
```{r mean, echo=TRUE}

 mean(totalStepsPerDay$steps)
 
```

Median of the total number of steps taken per day
```{r median, echo=TRUE}

 median(totalStepsPerDay$steps)
 
```

## What is the average daily activity pattern?

```{r dailyActivity, echo=TRUE}

library(ggplot2)

meanSteps <- aggregate(actDataNoNA$steps, list(interval = as.numeric(as.character(actDataNoNA$interval))), FUN = "mean")

names(meanSteps)[2] <- "avgNumberOfSteps"

ggplot(meanSteps, aes(interval, avgNumberOfSteps)) + geom_line(color = "grey", size = 0.6) + labs(title = "Average daily activity pattern", x = "5-minute intervals", y = "Average number of steps")

    
```  

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  

```{r maxNumOfSteps, echo=TRUE}

    meanSteps[meanSteps$avgNumberOfSteps == max(meanSteps$avgNumberOfSteps), ]

``` 

## Inputing missing values

* The total number of rows with NAs:

```{r numNA, echo=TRUE}

    sum(is.na(actData))

``` 

* Devise a strategy for filling in all of the missing values in the dataset. 

Using the mean for that 5-minute interval

```{r filledData, echo=TRUE}

  filledData <- actData 
  for (i in 1:nrow(filledData)) {
      if (is.na(filledData$steps[i])) {
          filledData$steps[i] <- meanSteps[which(filledData$interval[i] == meanSteps$interval), ]$avgNumberOfSteps
      }
  }
  
  # checking data correctness
  head(filledData)
  sum(is.na(filledData))  

``` 
* Histogram of the total number of steps taken each day 

 ```{r stepsPerDayFilledData, echo=TRUE}

ggplot(filledData, aes(date, steps)) + geom_bar(stat = "identity",colour = "grey",fill = "grey",width = 0.7) + labs(title = "Histogram of total number of steps whit filled na data", x = "Date", y = "Steps")

```  

* Mean of the total number of steps taken per day
```{r meanFilled, echo=TRUE}

   totalStepsPerDayFilled <- aggregate(steps ~ date, filledData, sum)
   mean(totalStepsPerDayFilled$steps)
 
```

* Median of the total number of steps taken per day
```{r medianFilled, echo=TRUE}

     median(totalStepsPerDayFilled$steps)
 
```
## Are there differences in activity patterns between weekdays and weekends?

```{r avgNumberOfStepsWk, echo=TRUE}
Sys.setlocale("LC_TIME", "English") 
                                   
filledData$weekdays <- weekdays(as.Date(filledData$date))
filledData$weekdays <- ifelse(filledData$weekdays %in% c("Saturday", "Sunday"),"weekend", "weekday")

meanStepsWk <- aggregate(filledData$steps, 
                      list(interval = as.numeric(as.character(filledData$interval)), 
                           weekdays = filledData$weekdays),
                      FUN = "mean")
                      
names(meanStepsWk)[3] <- "avgNumberOfStepsWk"    

library(lattice)

xyplot(avgNumberOfStepsWk ~ interval | weekdays,data= meanStepsWk,
       layout = c(1, 2), type = "l", col="grey", 
       xlab = "Interval", ylab = "Average number of steps")  
       
```       

    


